---
title: "p8105_hw1_baw2160"
author: "Bryce Wong"
date: "September 14, 2018"
output: github_document
---
``` {r loading_libraries, include = FALSE}

library(tidyverse)

```

# Problem One:

## Creating a data frame 

The prob1_df data frame contains one random sample of size 10 from a uniform[0, 5] distribution, one logical vector indicating whether elements of the sample are greater than 2, one character vector of length 10, and one factor vector of length 10:

```{r problem_1_data_frame}
prob1_df = tibble(
  unif_samp = runif(10, 0, 5),
  
  vec_logic = unif_samp > 2,
  
  vec_char = c("I", "just", "made", "this", "a", "character", "vector", "of", "size", "ten"),
  
  vec_factor = factor(c("dog", "cat", "cat", "dog", "dog", "dog", "bird", "cat", "dog", "cat"))
)
```

## Taking the mean
Attempting to take the mean of each variable in my date frame:
```{r problem_1_means}
mean(prob1_df[["unif_samp"]])
mean(prob1_df[["vec_logic"]])
mean(prob1_df[["vec_char"]])
mean(prob1_df[["vec_factor"]])
```
Turns out that taking the means of the character and factor vector do not work, but taking the means of the logical vector and the random sample from the uniform distribution do.

The error message indicates that one can only take the mean of a numeric or a logical vector when using the mean() function.

## Converting variables
In this code chunk, I attempt to convert my logical, character, and factor vector to a numerical vector:
```{r convert_to_num, results = "hide"}
as.numeric(prob1_df[["vec_logic"]])
as.numeric(prob1_df[["vec_char"]])
as.numeric(prob1_df[["vec_factor"]])
```
This works fine for the logical and factor vector, but it appears that I can't convert my character vector into a numerical vector.

In this next chunk, I attempt to convert my character variable to factor and then to numeric. I also try converting my factor variable from factor to character to numeric:
```{r convert_vec_char, results = "hide"}
vec_char.factor = factor(prob1_df[["vec_char"]])
as.numeric(vec_char.factor)
vec_factor.char = as.character(prob1_df[["vec_factor"]])
as.numeric(vec_factor.char)
```
Once the character variable was a factor variable, it could be converted to a numeric vector. But once the factor variable was turned into  a character variable, it couldn't be converted to a numeric vector. It looks like a character variable cannot become a numeric variable.

## Problem 2:

Now we are creating a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector indicating whether the x + y > 0
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r problem_2_dataframe}
prob2_df = tibble(
  x = rnorm(1000),
  
  y = rnorm(1000),
  
  vec_logic2 = x + y > 0,
  
  vec_num = as.numeric(vec_logic2),
  
  vec_factor2 = factor(vec_num)
)
```
The size of prob2_df is `r dim(prob2_df)`

The mean of x is `r mean(prob2_df$x)`

The median of x is `r median(prob2_df$x)`

The proportion of cases for which the logical vector is TRUE is `r mean(prob2_df$vec_logic2)`

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

Export your first scatterplot to your project directory using ggsave