---
title: "p8105_hw1_baw2160"
author: "Bryce Wong"
date: "September 14, 2018"
output: github_document
---
``` {r loading_libraries, include = FALSE}
library(tidyverse)
library(ggplot2)
```

# Problem One:

## Creating a data frame 

The prob1_df data frame contains one random sample of size 10 from a uniform[0, 5] distribution, one logical vector indicating whether elements of the sample are greater than 2, one character vector of length 10, and one factor vector of length 10:

```{r problem_1_data_frame}
prob1_df = tibble(
  unif_samp = runif(10, 0, 5),
  
  vec_logic = unif_samp > 2,
  
  vec_char = c("I", "just", "made", "this", "a", "character", "vector", "of", "size", "ten"),
  
  vec_factor = factor(c("dog", "cat", "cat", "dog", "dog", "dog", "bird", "cat", "dog", "cat"))
)
```

## Taking the mean
Attempting to take the mean of each variable in my date frame:
```{r problem_1_means}
mean(prob1_df[["unif_samp"]])
mean(prob1_df[["vec_logic"]])
mean(prob1_df[["vec_char"]])
mean(prob1_df[["vec_factor"]])
```
Turns out that taking the means of the character and factor vector do not work, but taking the means of the logical vector and the random sample from the uniform distribution do.

The error messages indicates that one can only take the mean of a numeric or a logical vector when using the mean() function.

This could be due to the fact that a character vector isn't made of values that can be summed and divided. Similarly, the factor vector is composed of categorical variables, which also can't be summed and divided like a continuous variable. Even though logical vectors could be technically be made of variable names that look like characters and represent dichotomous outcomes, they can be transformed into 1s and 0s, which can be evaluated like a continuous variable. 

## Converting variables
In this code chunk, I attempt to convert my logical, character, and factor vector to a numerical vector:
```{r convert_to_num, results = "hide"}
as.numeric(prob1_df[["vec_logic"]])
as.numeric(prob1_df[["vec_char"]])
as.numeric(prob1_df[["vec_factor"]])
```
This works fine for the logical and factor vector, but it appears that I can't convert my character vector into a numerical vector.

In this next chunk, I attempt to convert my character variable to factor and then to numeric. I also try converting my factor variable from factor to character to numeric:
```{r convert_vec_char, results = "hide"}
vec_char.factor = factor(prob1_df[["vec_char"]])
as.numeric(vec_char.factor)
vec_factor.char = as.character(prob1_df[["vec_factor"]])
as.numeric(vec_factor.char)
```
Once the character variable was a factor variable, it could be converted to a numeric vector. But once the factor variable was turned into  a character variable, it couldn't be converted to a numeric vector. It looks like a character variable cannot become a numeric variable.

# Problem 2:

## New dataframe:

Now we are creating a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution;
y: a random sample of size 1000 from a standard Normal distribution;
A logical vector indicating whether the x + y > 0;
A numeric vector created by coercing the above logical vector;
A factor vector created by coercing the above logical vector

```{r problem_2_dataframe}
prob2_df = tibble(
  x = rnorm(1000),
  
  y = rnorm(1000),
  
  vec_logic2 = (x + y) > 0,
  
  vec_num = as.numeric(vec_logic2),
  
  vec_factor2 = as.factor(vec_logic2)
)
```
## Describing the problem 2 dataframe:

The size of prob2_df is `r dim(prob2_df)`

The mean of x is `r mean(prob2_df$x)`

The median of x is `r median(prob2_df$x)`

The proportion of cases for which the logical vector is TRUE is `r mean(prob2_df$vec_logic2)`

## Making a scatterplot: 

```{r Making_scatterplots}
ggplot(prob2_df, aes(x = x, y = y, color = vec_logic2)) + geom_point()

ggsave(file = "C:/Users/Dell/Documents/3rd Semester - MPH/Data Science I/HW1/p8105_hw1_baw2160/first_scatplot.png")

ggplot(prob2_df, aes(x = x, y = y, color = vec_num)) + geom_point()

ggplot(prob2_df, aes(x = x, y = y, color = vec_factor2)) + geom_point()
```

It appears as if the color scale for the first scatterplot (using the logical vector) is the same as the color for the third scatterplot (using the factor vector). They are colored based on a binary labeling pattern, and the pattern appears exactly the same. This makes sense to me, because they both represent the same vector (the factor vector is created from the logical vector, so holds the same binary info - it just can't be evaluated as a continuous variable). 

The color scale for the second scatterplot (using the numeric vector) is a little different. In this one, the color can be anywhere along a gradient from dark to light, where the darkest point on the scale represents 0 and the lightest represents 1. In practice, however, the coloring pattern seems the same as the other two plots. Again, the data represented in the numeric vector is the same binary data represented in the logical and factor vector. So, there aren't any points that don't belong to either the darkest or lightest point in the scale, and the position of each point in the scatterplot is the same. 